---
title: "dc_arrests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

##Load libraries
```{r}
library(tidyverse)
#install.packages('lubridate')
library(lubridate)
#install.packages('janitor')
library(janitor)
```

download all data from 2020-2021, filter an review for dc area - Aadit
## Load conflict 


### Nicks Notes
- Is it supposed to be less than and greater than, or less than and equal etc...
- What are data ids? They are unique
- Does updated data have same numbers for the same dates?

```{r}

us_conflicts<- read_csv("data/us_conflicts.csv")

us_conflicts<- us_conflicts %>% 
  mutate(event_date = dmy(event_date))

jan_6_conflict<- us_conflicts %>% 
  filter(admin1 == "District of Columbia" & event_date == "2021-01-06") %>% 
  select(-data_id, -iso, -iso3, time_precision, -region, -country, -event_id_cnty, -event_id_no_cnty, -timestamp, -time_precision)
  
blm_conflict<- us_conflicts %>% 
  filter(admin1 == "District of Columbia" & event_date > "2020-05-30" & event_date < "2020-06-02") %>% 
  select(-data_id, -iso, -iso3, time_precision, -region, -country, -event_id_cnty, -event_id_no_cnty, -timestamp, -time_precision)

###############Nicks Check on data id column ####################


check_id = us_conflicts %>% 
  group_by(data_id) %>% 
  count()
```





## Load arrest data and rename columns so they're easier to work with

### Nicks Questions
- What are ccns? They are not unique, and it looks like they are location groups?
```{r}
dc_arrests<- read_csv("data/dc_arrest_data.csv")

dc_arrests<- dc_arrests %>% 
  clean_names()

```
mutate date columns to be dates not strings
```{r}
dc_arrests<- dc_arrests %>% 
   mutate(arrest_date = mdy(arrest_date),
          offense_date = mdy(offense_date))


### Nicks Checking ######
ccn_check  = dc_arrests %>% 
  group_by(ccn) %>% 
  count()
```
data starts at 5-30-20, main BLM protests through 6-3-20, 5-31 curfew was put in place, ended 6-3

5-31 - church bible 

insurrection 1-6-21

## Questions

Questions to ask:

1. what does unrest related data mean?
"individuals arrested and processed for curfew violations, riot-related burglary, and other riot-related events" from (DC Metro PD [https://mpdc.dc.gov/publication/mpd-unrest-related-arrest-data-set])

right ran rallies- do more people get killed?

talk to professors that teach conflict related stuff/people who study conflict? what causes conflict and civil unrest?

##Data viz potential: 
1. How many people were arrested at each event?
  - compare individual days of protests to insurrection
  
### Nick Data Check
- If main blm protests went through 06-3-20, arent the date filters in the code wrong??
- The same person can show up more than one time .....
  - Jan 6th dc arrests joshua pruitt shows up twice for example
- Jan 6th offenses you filtered for arrest instead of offenses like you did in blm

```{r}
#blm arrests/offenses
blm_arrests<- dc_arrests %>% 
  filter(arrest_date > "2020-05-30" & arrest_date < "2020-06-02") %>% 
  group_by(arrest_date) %>% 
  summarise(
    arrest_count = n()
  )

blm_offenses<- dc_arrests %>% 
  filter(offense_date > "2020-05-30" & offense_date < "2020-06-02") %>% 
  group_by(offense_date) %>% 
  summarise(
    offense_count = n()
  )






blm_arrest_info <- blm_arrests %>% 
  left_join(blm_offenses, by=c('arrest_date'='offense_date')) %>% 
  rename(date = arrest_date)
# skipping 6-4 for some reason???

#insurrection arrests/offenses

jan6_arrests<- dc_arrests %>% 
  filter(arrest_date > "2021-01-05" & arrest_date < "2021-01-07") %>% 
  group_by(arrest_date) %>% 
  summarise(
    arrest_count = n()
  )

jan6_offenses<- dc_arrests %>% 
  filter(arrest_date > "2021-01-05" & arrest_date < "2021-01-07") %>% 
  group_by(offense_date) %>% 
  summarise(
    offense_count = n()
  )

jan6_arrest_info <- jan6_arrests %>% 
  left_join(jan6_offenses, by=c('arrest_date'='offense_date')) %>% 
  rename(date = arrest_date)

################################# Nick Checking uniqueness of names for blm ################################

blm_dc_arrest_name = dc_arrests %>%
  filter(arrest_date > "2020-05-30" & arrest_date < "2020-06-02") %>% 
  group_by(defendant_first_name, defendant_last_name, age_at_arrest ) %>% 
  count()

blm_dc_offense_name = dc_arrests %>%
  filter(offense_date > "2020-05-30" & offense_date < "2020-06-02") %>% 
  group_by(defendant_first_name, defendant_last_name, age_at_arrest ) %>% 
  count()

blm_dc_arrest_name_all = dc_arrests %>%
  group_by(defendant_first_name, defendant_last_name, age_at_arrest ) %>% 
  count()

#####


##################################Nick Checking Uniqueness of names for riots ######################################
jan6_arrest_name<- dc_arrests %>% 
  filter(arrest_date > "2021-01-05" & arrest_date < "2021-01-07") %>% 
  group_by(defendant_first_name, defendant_last_name, age_at_arrest ) %>% 
  count()


jan6_offense_name<- dc_arrests %>% 
  filter(offense_date > "2021-01-05" & offense_date < "2021-01-07") %>% 
  group_by(defendant_first_name, defendant_last_name, age_at_arrest ) %>% 
  count()

jan6_arrests_ungrouped <- dc_arrests %>% 
  filter(arrest_date > "2021-01-05" & arrest_date < "2021-01-07")

jan6_offense_ungrouped <- dc_arrests %>% 
  filter(offense_date > "2021-01-05" & offense_date < "2021-01-07")

```
  
Make df to create graphic for number of arrests (based on offense date)
```{r}

arrests_by_offense_date<- bind_rows(jan6_arrest_info, blm_arrest_info) %>% 
  select(-arrest_count)
  
#write_csv here
write_csv(arrests_by_offense_date, "arrests_by_offense_date")

```


  
2. What were the most common charges for each event?

### Nick questions
- Why do you start filtering on 05-29 when you started on 05_30 for previous code?

```{r}
#blm charges
blm_charges<- dc_arrests %>% 
  filter(offense_date > "2020-05-29" & offense_date < "2020-06-06") %>% 
  group_by(charge_s, offense_date) %>% 
  summarise(
   count=n()
  ) %>% 
  arrange(desc(count)) 
  #head(10)

blm_charges_split<- blm_charges %>% 
  mutate(charge_s = strsplit(str_squish(as.character(charge_s)), ",")) %>% 
  unnest(charge_s) %>% 
  mutate(charge_s = str_trim(charge_s,side="both")) %>% 
  group_by(charge_s, offense_date) %>% 
  summarise(
   sum_arrests = sum(count)
  )

#blm_may_31
blm_may_31<- blm_charges_split %>% 
  filter(offense_date == "2020-05-31") %>% 
  arrange(desc(sum_arrests)) 
  #%>% 
  #group_by(charge_s, arrest_date) %>% 
  #summarise(
   #count=n()
  #) %>% 
  #arrange(desc(count)) 
  #head(10)

#blm_june_1
blm_june_1<- blm_charges_split %>% 
  filter(offense_date == "2020-06-01")%>% 
  arrange(desc(sum_arrests)) 
  #%>% 
  #group_by(charge_s, arrest_date) %>% 
  #summarise(
   #count=n()
  #) %>% 
  #arrange(desc(count))
  #head(10)


```

```{r}

#jan6_charges
jan6_charges<- dc_arrests %>% 
  filter(offense_date == "2021-01-06") %>%  
  group_by(charge_s, offense_date) %>% 
  summarise(
   count=n()
  ) %>% 
  arrange(desc(count)) 
  #head(10)

jan6_charges_split<- jan6_charges %>% 
  mutate(charge_s = strsplit(str_squish(as.character(charge_s)), ",")) %>% 
  unnest(charge_s) %>% 
  mutate(charge_s = str_trim(charge_s,side="both")) %>%
  mutate(charge_s = case_when(
    charge_s == "Unlawful Entry" ~ "Unlawful entry",
    #this "replaces" existing data that's spelled correctly with itself
    TRUE ~ charge_s
  )) %>%
  group_by(charge_s, offense_date) %>% 
  summarise(
   sum_arrests = sum(count)
  )%>% 
  arrange(desc(sum_arrests))

```
Exporting top 5 charges for each event 
```{r}

blm_531_charges<- blm_may_31 %>% 
  head(5)

blm_601_charges<- blm_june_1 %>% 
  head(5)

jan6_top_charges<- jan6_charges_split %>% 
  head(5)

write.csv(blm_531_charges, "blm_531_charges.csv")
write.csv(blm_601_charges, "blm_601_charges.csv")
write.csv(jan6_top_charges, "jan6_top_charges.csv")

```

Also doing one big list to see if this chart would look cooler
**IT DIDN"T**
```{r}

top_charges<- bind_rows(blm_531_charges,blm_601_charges,jan6_top_charges)

write.csv(top_charges, "top_charges.csv")


```
3. How many people were arrested specifically for curfew violations? 

```{r}
blm_curfew_531<- blm_may_31%>% 
  filter(charge_s == "Curfew Violation")

blm_curfew_601<- blm_june_1%>% 
  filter(charge_s == "Curfew Violation")

jan6_curfew<- jan6_charges_split %>% 
  filter(charge_s == "Curfew Violation")

curfew_violations<- bind_rows(blm_curfew_531, blm_curfew_601, jan6_curfew)

#write_csv here
write_csv(curfew_violations, "curfew_violations")
```


### Data Sentences Nick Check

arrested 363 people over two days of Black Lives Matter demonstrations

```{r}

```

only 61 people at the Jan. 6 insurrection

```{r}

```

Black Lives Matter protestors were arrested over five times more than people who took part in the Jan. 6 insurrection.

```{r}

```


Curfew Violations: Insurrection Arrests Were a Fraction of Black Lives Matter Arrests

```{r}

```


The difference between curfew violation arrests between the two events is also stark, with five times more arrests during the Black Lives Matter demonstrations than the Capitol demonstration.

```{r}

```

Apart from curfew violations, Black Lives Matter demonstrations saw the highest concentration of riot act- felony and second-degree burglary charges

```{r}

```

On Jan. 6, unlawful entry was the second most common charge. There was only one arrest for a riot act- felony on Jan. 6. 

```{r}

```

