---
title: "dc_arrests"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Load libraries
```{r}
library(tidyverse)
#install.packages('lubridate')
library(lubridate)
```


Load data and rename columns so they're easier to work with
```{r}
dc_arrests<- read_csv("data/dc_arrest_data.csv")

dc_arrests<- dc_arrests %>% 
  rename("arrest_date" = "Arrest Date",
         "offense_date" = "Offense Date",
         "offense_block_location" = "Offense Block Location",
         "first_name" = "Defendant First Name",
         "last_name"= "Defendant Last Name",
         "age_at_arrest"="Age at Arrest",
         "state" = "State",
         "charge" = "Charge(s)"
         )

```
mutate date columns to be dates not strings
```{r}
dc_arrests<- dc_arrests %>% 
   mutate(arrest_date = mdy(arrest_date),
          offense_date = mdy(offense_date))
```
data starts at 5-30-20, main BLM protests thru 6-3-20, 5-31 curfew was put in place, ended 6-3

5-31 - church bible 

insurrection 1-6-21

Questions to ask:
what does unrest related data mean?
right rallies- do more people get killed?
download all data from 2020-2021 in dc area - Aadit
talk to professors that teach conflict related stuff? people who study conflict? what causes conflict and civil unrest?
1. How many people were arrested at each event?
  - compare individual days of protests to insurrection
2. How many people were arrested specifically for curfew violations? 
3. How many minors were arrested at each event?
4. What were the most common charges for each event?


